---
title: "Anova"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Between groups {#anchor}  
**Application** : Comparing means of a quantitative variable obtained from two independent groups.  
**Research Hypothesis** : The researcher hypothesized that students who eat salty food exercise more than their sweet and savory counterparts.  
**Null Research Hypothesis** : There is no difference in the reported amounts of exercise for salty, savory, and sweet people.  

**First load packages and library**  
> point and click on the folder under the Environment Tab
```{r echo = FALSE}
class <- read.csv("C:/Users/Branly Mclanbry/Dropbox/Psych 242 ADD Health Codebook/Fall 2017/R/example.csv",  na.strings="999")
```
**Install packages if you have not installed them**
```{r eval = FALSE}
install.packages(c("tidyverse", "lsr", "multcomp"))
```
**Load them**
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lsr)
library(multcomp)
```

Factor levels for `flavor` because '1' flavor does not make sense. Recode variables into a new variable instead of overwriting the original codes.
Now codes like '1' are salty while '2' is sweet.
```{r}
class$flavortown[class$Flavor==1] <- "salty"
class$flavortown[class$Flavor==2] <- "sweet"
class$flavortown[class$Flavor==3] <- "savory"
```
Alternatively, another `recode_factor` function comes from the `tidyverse` package and expedites the factoring process. 
```{r}
class$flavortown <- recode_factor(class$Flavor, `1` = "salty", `2` = "sweet", `3` = "savory")
```
**Analysis**  
Next is the creation of the actual model with `flavor` against `exercise` and the data is from the `class` datafile.  
```{r}
flavorlandaov <- aov(exercise~flavortown, data = class)
```
**Results**  
We want to find the results of the analysis of variance.  
```{r}
summary(flavorlandaov)
```
Our analysis has revead that the *p*-value is .03 which generally means that there is a 3% chance of a Type 2 error.   
  
**Tukey tests**  
Now that we know there are a difference between the groups of `flavor`, we have to find out which groups actually excercise more. a *Tukey* test is an analysis that provides a more in-depth look at the differences between the means while dictating which relations are statistically significant or not.  
The `glht` function comes from the `multcomp` package. This is a general linear hypothesis test.  
**Analysis**
```{r}
tukey <- glht(flavorlandaov,linfct=mcp(flavortown = "Tukey"))
```
**Results**
```{r}
summary(tukey)
```
We can look at the *p*-values which tells us that there is a difference between reported exercise amounts of the `sweet-salty` and the `savory-salt` relationship.  
  
**Information for write up**  
We will need some more information for the write up, specifically the sample size, means, standard deviations and effect size.  
**Means and Standard Deviation**  
This can be done with the `tapply` function. The command requests that the `sd` or `mean` function be applied to `class$exercise` and `class$flavor`.
```{r}
tapply(class$exercise,class$flavortown, sd)
tapply(class$exercise,class$flavortown, mean)
```
**Sample size**   
Then we will need to know the sample size per group. This will be accomplished with the `table` function.
```{r}
table(class$flavortown)
```
**Effect size**   
Effect sizes can be found from the `lsr` package utilizing the `etaSquared` function.
```{r}
etaSquared(flavorlandaov)
```
**Write up**  
There is a difference between flavor preference and reported exercise hours, *F*(2,38) = 3.73, *p* < .05, partial n^2^ = .16. salty (*M* = 18.00, *SD* = 16.10) excercised more than their Sweet (*M* = 7.28, *SD* = 5.33), *p* < .05 and savory (*M* = 7.37, *SD* = 6.82), *p* < .05 counterparts. No differences were found between the sweet and savory parts, *p* = .99. 



| Table 1          |       |       |       |
|--------------------|-------|-------|-------|
| *Summary Statistics* |       |       |       |
| Variable           | *M*     | *SD*    | n     |
| Salty              | 18.00 | 16.10 | 4     |
| Sweet              | 7.28  | 5.33  | 18.00 |
| Savory             | 7.37  | 6.81  | 19.00 |


**Plots**  
Usually information can be provided with plots which can easy cognitive loads by visually presenting information.  
This is done with the `barplot` function.  
First a model of what you want to plot has to be created. I named my model `saltyfitness`.
```{r}
saltyfitness <- tapply(class$exercise,class$flavortown, mean)
```
Then `saltyfitness` can be wrapped with a barplot. Further parameters can be added, such as `xlab` which is a label for the axis or `ylab` which is the y axis label. Things like color or border can be added as well.
```{r}
barplot(saltyfitness, xlab = "Flavor", ylab = "reported hours exercising", col = c("maroon1","springgreen1","yellow"),border = "navyblue" )
```

